Working title: Highly resolved spatiotemporal simulations for exploring mixed fishery dynamics

### 2019-10-06 write_up/mixfishsim_ecolmod.tex
Response to JJP comments

We parametrised the simulation model for four example populations with
different demographics, growth rates, natural mortality and recruitment
parameters (Table \ref{tab:1}). Habitat preference (Figure \ref{S-fig:1}) and
temperature tolerances (Figures \ref{S-fig:3}, \ref{S-fig:4}) were defined to
be unique to each population resulting in differently weekly distribution
patterns (Figures \ref{S-fig:5}-\ref{S-fig:7}). In addition, each of the
populations was assumed to have two defined spawning areas that result in the
populations moving towards these areas in pre-defined weeks (Figure
\ref{S-fig:2}) with population-specific movement rates (Table \ref{tab:1}).

@@

Survey closures were on an annual basis only, as this was the most temporally
resolved survey data available.

@@

Different perceptions of the proportion of each stock in an area are seen when
we aggregate the data over a smaller geographical region at different
timescales, with weekly (top), monthly (middle) and yearly (bottom) catch
compositions from across an aggregated 20 x 20 area (Figure \ref{fig:2}). In
the true population, the monthly aggregation captures the major patterns of
composition seen in the weekly data with the percentage of different
populations in the catch having similar mean and standard deviations.  In the
weekly data population 1 = 9.36 (3.99), population 2 = 83.2 (5.60), population
3 = 3.57 (1.23), population 4 = 3.91 (1.59); in the monthly data population 1 =
9.23 (3.87), population 2 = 83.3 (5.52), population 3 = 3.62 (1.15), population
4 = 3.86 (1.52). While means were similar some of the variation was lost when
aggregated to an annual level; population 1 = 9.90 (0.173), population 2 = 82.2
(0.308), population 3 = 3.82 (0.119), population 4 = 4.03 (0.0502).\\

The commercial data on a weekly basis shows some of the same patterns as the
true population, though the population 1 (in red) is less well represented and
some weeks are missing catches from the area. Here, weekly and monthly
compositions were nearly identical, with monthly composition of population 1 =
0.0472 (0.0139), population 2 = 94.4 (1.47), population 3 = 3.12 (1.47),
population 4 = 2.40 (0.444). Again, yearly values head a similar mean but
smaller standard deviation.\\

The survey data was only available on an annual basis, and showed again a
slightly different composition; population 1 = 0.372 (0.00473), population 2 =
87.7 (0.193), population 3 = 0.729 (0.0200), population 4 = 11.2 (0.172).

@@

The trend in fishing mortality for each species show that in most cases the
fishery closure was successful in reducing fishing mortality on the species of
interest (population 3; Figure \ref{fig:3}), though interestingly the largest
reductions in fishing mortality happened immediately after the closures,
following which the fisheries ``adapted" to the closures and fishing mortality
increased again somewhat. The exception to the success was the closures
implemented based on the coarsest spatial (20 x 20) and temporal resolution
(yearly) that was ineffective with all data sources. As expected, closures
based on the ``known" population distribution were most effective, with
differing degrees of success using the commercial data. Fishing mortality rates
on the other species changed in different proportions, depending on whether the
displaced fishing effort moved to areas where the populations were found in
greater or lesser density. \\

A regression tree (using the R package REEMtree \citep{Sela2012}) highlights
that the factor most contributing to differences in fishing mortality before
and after the closure was the population (72 \% showing that the closures were
effective for population 3), followed by data resolution (21 \%), data type (7
\%) with the least important factor the timescale ($<$ 1 \%). In general the
finer the spatial resolution of the data used the greater reduction in fishing
mortality for population 3 after the closures (Figure \ref{fig:4}). The notable
outliers are the commercial data at the coarsest spatial resolution (20 x 20)
at a yearly and weekly timescale, where closures were nearly as effective as
the fine-scale resolution. In this case the closures were sufficiently large to
protect a core area of the habitat for the population, but this was achieved in
a fairly crude manner by closing a large area - including area where the
species was not found (Figure \ref{fig:17}) that may have consequences in terms
of restricting the fishery in a much larger area than necessary. \\

@@

%% General intro
Our study presents a new highly resolved fisheries simulation framework,
evaluates the importance of data scaling and considers potential bias
introduced through data aggregation when using fisheries data to infer
spatiotemporal dynamics of fish populations. Understanding how fishers exploit
multiple heterogeneously distributed fish populations with different catch
limits or conservation status requires detailed understanding of the overlap of
resources; this is difficult to achieve using conventional modelling approaches
due to species targeting in fisheries resulting in preferential sampling
\citep{Martinez-Minaya2018}. Often data are aggregated or extrapolated which
requires assumptions about the spatial and temporal scale of processes. Our
study explores the assumptions behind such aggregation and preferential
sampling to identify potential impacts on management advice. With modern
management approaches increasingly employing more nuanced spatiotemporal
approaches to maximise productivity while taking account of both the biological
and human processes operating on different time-frames \citep{Dunn2016},
understanding assumptions behind the data used - increasingly a combination of
logbook and positional information from vessel monitoring systems - is vital to
ensure measures are effective. \\

@@

Two contrasting real world approaches in this respect were the spatial closures
to protect cod in the North Sea. In one example, large scale spatial closures
were implemented with little success due to effort displacement to previously
unfished areas \citep{Dinmore2003}, while in another small scale targeted
spatiotemporal closures were considered to have some effect in reducing cod
mortality without having to disrupt other fisheries significantly
\citep{Needle2011}. These examples emphasise the importance of considering the
right scale and aggregation of data when identifying area closures and the need
to consider changing dynamics in the fisheries in response to such closures. \\

@@

We model the population and fleet dynamic processes to draw inference on the
importance of data scale and aggregation in understanding and managing mixed
fisheries and their impact on multiple fish populations. In doing so, we have
necessarily had to make a number of simplifying assumptions. \\

@@

\begin{figure}[!ht]
	\includegraphics[width = \linewidth]{Plots/Combined_Movement}
	\caption{(A) The fishing locations (points) and movements (lines) of a
		single vessel during a trip overlaid on the revenue of a
		fishing site (landings x price); (B) the fishing locations of
		the vessel over several trips (value field changes over the
		period so not shown). Note that movements are a mixture of
		correlated random walk (solid lines) and experience-based
		(dashed lines), and that the field is wrapped on a torus so
		that opposite sides of the spatial domain are considered next
		door; (C) the locations of multiple vessels from the
		same fleet overlaid on the value field, (D) the realised step
		distance and turning angles for a single vessel over the
		simulation.}
	\label{fig:mv}
\end{figure}

>>>>

We parametrised the simulation model for four example populations with
different demographics, growth rates, natural mortality and recruitment
parameters (Table \ref{tab:1}). Habitat preference (Figure \ref{S-fig:1}) and
temperature (Figures \ref{S-fig:3}, with temperature tolerance \ref{S-fig:4})
defined to be unique to each population resulting in differently weekly
distribution patterns (Figures \ref{S-fig:5}-\ref{S-fig:7}). In addition, each
of the populations was assumed to have two defined spawning areas that result
in the populations moving towards these areas in pre-defined weeks (Figure
\ref{S-fig:2}) with population-specific movement rates (Table \ref{tab:1}).

@@

Survey closures were on an annual basis only, as this was the most temporally
resolved survey data available. We evaluated the factors contributing to the
success of the closures through a regression tree (using the R package REEMtree
\citep{Sela2012}) to identify the factor most contributing to differences in
fishing mortality before and after the closure.

@@

Different perceptions of the proportion of each stock in an area are seen when
we aggregate the data over a smaller geographical region at different
timescales, with weekly (top), monthly (middle) and yearly (bottom) catch
compositions from across an aggregated 20 x 20 area (Figure \ref{fig:2}). In
the true population, the monthly aggregation captures the major patterns of
composition seen in the weekly data with the percentage of different
populations in the catch having similar mean and standard deviations.  In the
weekly data population 1 = 9.36\% (SD = 3.99), population 2 = 83.2\% (5.60),
population 3 = 3.57\% (1.23), population 4 = 3.91\% (1.59); in the monthly data
population 1 = 9.23\% (3.87), population 2\% = 83.3 (5.52), population 3 =
3.62\% (1.15), population 4 = \%3.86 (1.52). While means were similar some of
the variation was lost when aggregated to an annual level; population 1 =
9.90\% (0.173), population 2 = 82.2\% (0.308), population 3 = 3.82\% (0.119),
population 4 = 4.03\% (0.0502).\\

The commercial data on a weekly basis shows some of the same patterns as the
true population, though the population 1 (in red) is less well represented and
some weeks are missing catches from the area. Here, weekly and monthly
compositions were nearly identical, with monthly composition of population 1 =
0.0472\% (0.0139), population 2 = 94.4\% (1.47), population 3 = 3.12\% (1.47),
population 4 = 2.40\% (0.444). Again, yearly values head a similar mean but
smaller standard deviation.\\

The survey data was only available on an annual basis, and showed again a
slightly different composition; population 1 = 0.372\% (0.00473), population 2
= 87.7\% (0.193), population 3 = 0.729\% (0.0200), population 4 = 11.2\%
(0.172).

@@

In most cases the fishery closure was successful in reducing fishing mortality
on the species of interest (population 3; Figure \ref{fig:3}), though
interestingly the largest reductions in fishing mortality happened immediately
after the closures, following which the fisheries ``adapted" to the closures
and fishing mortality increased again somewhat. The exception to the success
was the closures implemented based on the coarsest spatial (20 x 20) and
temporal resolution (yearly) that was ineffective (i.e. failed to reduce
fishing mortality) with all data sources. As expected, closures based on the
``known" population distribution were most effective, with differing degrees of
success using the commercial data. Fishing mortality rates on the other species
changed in different proportions, depending on whether the displaced fishing
effort moved to areas where the populations were found in greater or lesser
density. \\

The factor most contributing to differences in fishing mortality before and
after the closure was the population (72 \% showing that the closures were
effective for population 3), followed by data resolution (21 \%), data type (7
\%) with the least important factor the timescale ($<$ 1 \%). In general the
finer the spatial resolution of the data used the greater reduction in fishing
mortality for population 3 after the closures (Figure \ref{fig:4}). The notable
outliers are the commercial data at the coarsest spatial resolution (20 x 20)
at a yearly and weekly timescale, where closures were nearly as effective as
the fine-scale resolution. In this case the closures were sufficiently large to
protect a core area of the habitat for the population, but this was achieved in
a fairly crude manner by closing a large area - including area where the
species was not found (Figure \ref{fig:17}) that may have consequences in terms
of restricting the fishery in a much larger area than necessary. \\

@@

%% General intro
Our study presents a new highly resolved fisheries simulation framework to
evaluate the importance of data scaling and considers potential bias introduced
through data aggregation when using fisheries data to infer spatiotemporal
dynamics of fish populations. Understanding how fishers exploit multiple
heterogeneously distributed fish populations with different catch limits or
conservation status requires detailed understanding of the overlap of
resources; this is difficult to achieve using conventional modelling approaches
due to species targeting in fisheries resulting in preferential sampling
\citep{Martinez-Minaya2018}. Often data are aggregated or extrapolated which
requires assumptions about the spatial and temporal scale of processes. Our
study explores the assumptions behind such aggregation and preferential
sampling to identify potential impacts on management advice. With modern
management approaches increasingly employing more nuanced spatiotemporal
approaches to maximise productivity while taking account of both the biological
and human processes operating on different time-frames \citep{Dunn2016},
understanding assumptions behind the data used - increasingly a combination of
logbook and positional information from vessel monitoring systems - is vital to
ensure measures are effective. \\

@@

Two contrasting real world approaches in this respect were the spatial closures
to protect cod in the North Sea. In one example, large scale spatial closures
were implemented with little success due to effort displacement to previously
unfished areas \citep{Dinmore2003}, while in another small scale targeted
spatiotemporal closures were considered to have some effect in reducing cod
mortality without having to disrupt other fisheries substantially
\citep{Needle2011}. These examples emphasise the importance of considering the
right scale and aggregation of data when identifying area closures and the need
to consider changing dynamics in the fisheries in response to such closures. \\

@@

We modelled the population and fleet dynamic processes to draw inference on the
importance of data scale and aggregation in understanding and managing mixed
fisheries and their impact on multiple fish populations. In doing so, we
necessarily had to make a number of simplifying assumptions. \\

@@

\begin{figure}[!ht]
	\includegraphics[width = \linewidth]{Plots/Combined_Movement}
	\caption{(A) The fishing locations (points) and movements (lines) of a
		single vessel during a trip overlaid on the revenue of a
		fishing site (landings x price); (B) the fishing locations of
		the vessel over several trips (value field changes over the
		period so not shown). Note that movements are a mixture of
		correlated random walk (solid lines) and experience-based
		(dashed lines), and that the field is wrapped on a torus so
		that opposite sides of the spatial domain are considered
		spatially close; (C) the locations of multiple vessels from the
		same fleet overlaid on the value field, (D) the realised step
		distance and turning angles for a single vessel over the
		simulation.}
	\label{fig:mv}
\end{figure}
